Capstone Movielens Report

### 2025-05-17 reports/capstone-movielens.site/02.5.UME-model.Rmd
RMD report updated: `data.plot` function description completed.
RMD Report update: UME Model Regularization Metod described.

@@

::: {.noteblock data-latex=""}
We use the custom data visualization function [data.plot](#func.data.plot) described in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report, which is defined in the [Data Visualization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L447) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L592) script.
:::

@@

#### User+Movie Effect Model: Fine-Tuning

>>>>

  3. **Retraining:** retraining the model with the best value of the parameter $\lambda$ obtained in the previous step.

@@

::: {.noteblock data-latex=""}
We use the custom data visualization function [data.plot](#func.data.plot) described in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report, which is defined in the [Data Visualization](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L447) section of the [data.helper.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L592) script on _GitHub_.
:::

@@

#### User+Movie Effect Model: Fine-Tuning
\

We are now ready to perform the fine-tuning step of our model _regularization_ process to determine the best value for the $\lambda$ parameter.

Here we are going to use the [model.tune.param_range](#func.model.tune.param_range) function described in the [Regularization: Common Helper Functions](#appndx_a.rglr.common_helper.functions) section of the [Appendix] to this report, passing the [regularize.test_lambda.UM_effect.cv](#func.regularize.test_lambda.UM_effect.cv) function as the value of the [fn_tune.test.param_value](#func.model.tune.param_range.params.fn_tune.test.param_value) parameter:
```{r, eval=FALSE}
endpoints <- 
  get_fine_tune.param.endpoints(cv.UME.preset.result$tuned.result)

UM_effect.loop_starter <- c(endpoints["start"], 
                            endpoints["end"], 
                            8)
UM_effect.loop_starter
#> [1] 0.3 0.5 8.0

UME.rglr.fine_tune.cache.base_name <- "UME.rglr.fine-tune"

UME.rglr.fine_tune.results <- 
  model.tune.param_range(UM_effect.loop_starter,
                         UME.rglr.fine_tune.cache.path,
                         UME.rglr.fine_tune.cache.base_name,
                         regularize.test_lambda.UM_effect.cv)
                         #endpoint.min_diff = 1e-07/4)

UME.rglr.fine_tune.RMSE.best <- UME.rglr.fine_tune.results$best_result["best_RMSE"]
```
```{r}
writeLines("*** Fine-tuning results object data structure ***")
str(UME.rglr.fine_tune.results)
writeLines("--------")
writeLines('')

writeLines("*** Fine-tuning: best results ***")
UME.rglr.fine_tune.results$best_result
writeLines("------")
writeLines('')

writeLines("*** Fine-tuning: best parameter value ***")
UME.rglr.fine_tune.RMSE.best
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section can be also found in the [Fine-tuning Step of the Regularization Method for the User+Movie Model](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L867) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R) script on _GitHub_.
:::

Let's visualize the fine-tuning results:
```{r}
UME.rglr.fine_tune.results$tuned.result |>
  data.plot(title = "Fine-tune Stage results of the Regularization Process for the User+Movie Model",
              xname = "parameter.value",
              yname = "RMSE",
              xlabel = TeX(r'[$\lambda$]'),
              ylabel = str_glue("Deviation from the best RMSE value (",
                                as.character(round(UME.rglr.fine_tune.RMSE.best, digits = 7)), 
                                ")"),
              normalize = TRUE)
```
\newpage

#### Retraining User+Movie Effect Model with the best $\lambda$
\

Finally, we retrain our model with the best $\lambda$ parameter value we have just figured out:
```{r, eval=FALSE}
  UME.rglr.best_lambda <- best_result["param.best_value"]
  UME.rglr.best_lambda

  rglr.UM_effect <- train_user_movie_effect(edx, UME.rglr.best_lambda)

```
```{r}
writeLines("*** The Best Fine-tuning Results ***")
UME.rglr.fine_tune.results$best_result
writeLines("------")
writeLines('')

  print_log1("Re-training Regularized User+Movie Effect Model for the best `lambda`: %1...",
           UME.rglr.best_lambda)
writeLines("------")
writeLines('')

writeLines("*** Regularized User+Movie Effect Structure ***")
str(rglr.UM_effect)
writeLines("------")
writeLines('')
```

::: {.noteblock data-latex=""}
The complete version of the source code provided in this section are available in the [Re-training Regularized User+Movie Effect Model for the best $\lambda$](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L910) section of the [capstone-movielens.main.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/capstone-movielens.main.R#L910) script on _GitHub_.
:::

Finally, after _Regularization_, we obtained the following result:
```{r, eval=FALSE}
UME.rglr.retrain.RMSE <- calc_user_movie_effect_RMSE.cv(rglr.UM_effect)

RMSEs.ResultTibble.rglr.UME <- RMSEs.ResultTibble.UME |> 
  RMSEs.AddRow("Regularized User+Movie Effect Model", 
               UME.rglr.retrain.RMSE,
               comment = "Computed for `lambda` = %1" |>
                 msg.glue(UME.rglr.best_lambda))
```
```{r}
print_log1("Regularized User+Movie Effect Model has been re-trained for the best `lambda`: %1.",
         UME.rglr.best_lambda)
print_log1("The best RMSE after being regularized: %1",
         UME.rglr.retrain.RMSE)

RMSE_kable(RMSEs.ResultTibble.rglr.UME)
```
